A Jacobian-free Newton-Krylov method for cell-centred finite volume solid mechanics

### 2025-09-19 article.tex
Add description of and comment on left preconditioning

%The literature indicates that the choice of preconditioner for the inner linearised system has a major impact on the efficiency and robustness of the overall solution procedure.
The purpose of preconditioning the Jacobian-free Newton-Krylov method is to reduce the number of inner linear solver iterations.
The current work uses the GMRES linear solver for the inner system.
%Left or right preconditioning, may be employed in a Jacobian-free context, and there are pros and cons to both.
Using right preconditioning, the finite difference approximation of Equation \ref{eq:JF} becomes
\begin{eqnarray}
%	(\bb{J} \bb{P}^{-1}) (\bb{P} \delta \bb{u}) = -\bb{R}(\bb{u})
	\bb{J} \bb{P}^{-1} \bb{v}
	\approx
	\frac{\bb{R}(\bb{u} + \epsilon \bb{P}^{-1} \bb{v}) - \bb{R}(\bb{u})}{\epsilon}
\end{eqnarray}
where $\bb{P}$ is the preconditioning matrix or process.
In practice, only the action of $\bb{P}^{-1}$ on a vector is required, and $\bb{P}^{-1}$ need not be explicitly formed.
Concretely, the preconditioner needs to approximately solve the linear system $\bb{y} = \bb{P}^{-1} \bb{v}$.
%Thus, while we may refer to the matrix P, operationally the algorithm only requires the action of $P^{-1}$ on a vector.

>>>>

%The literature indicates that the choice of preconditioner for the inner linearised system has a major impact on the efficiency and robustness of the overall solution procedure.
The purpose of preconditioning the Jacobian-free Newton-Krylov method is to reduce the number of inner linear solver iterations.
The current work uses the GMRES linear solver for the inner system.
%Left or right preconditioning, may be employed in a Jacobian-free context, and there are pros and cons to both.
Using right preconditioning, the finite difference approximation of Equation \ref{eq:JF} becomes
\begin{eqnarray}
%	(\bb{J} \bb{P}^{-1}) (\bb{P} \delta \bb{u}) = -\bb{R}(\bb{u})
	\bb{J} \bb{P}^{-1} \bb{v}
	\approx
	\frac{\bb{R}(\bb{u} + \epsilon \bb{P}^{-1} \bb{v}) - \bb{R}(\bb{u})}{\epsilon}
\end{eqnarray}
where $\bb{P}$ is the preconditioning matrix or process.
In practice, only the action of $\bb{P}^{-1}$ on a vector is required, and $\bb{P}^{-1}$ need not be explicitly formed.
Concretely, the preconditioner needs to approximately solve the linear system $\bb{y} = \bb{P}^{-1} \bb{v}$.
%Thus, while we may refer to the matrix P, operationally the algorithm only requires the action of $P^{-1}$ on a vector.
Alternatively, left preconditioning may be applied. In this case, the finite difference approximation of Equation \ref{eq:JF} becomes
\begin{eqnarray}
	\bb{P}^{-1} \bb{J} \bb{v} \;\approx\; \frac{\bb{P}^{-1}\left[ \bb{R}(\bb{u}+\epsilon \bb{v}) - \bb{R}(\bb{u}) \right]}{\epsilon}
\end{eqnarray}
where the preconditioner acts directly on the residual evaluations. As with right preconditioning, the explicit inverse of $\bb{P}$ is not required; only the ability to approximately apply $\bb{P}^{-1}$ to a given vector is necessary.
In practice, the choice between left and right preconditioning in Jacobian-free Newton–Krylov methods is not always straightforward. Right preconditioning preserves the true nonlinear residual for convergence checks, while left preconditioning alters the residual norm through the preconditioner. In this work we use left preconditioning, but no significant difference in convergence behaviour was observed between the two approaches.
